### 2024-01-18 meetings/18-01-2024_slides.qmd
slides meeting 18-01-2024

---
title: "Meeting --- Summary LLM"
author: "Bas Doners, Dennis Fok, Finn-Ole Höner"
date: 2024-01-18
date-format: long
format: beamer
aspectratio: 169
incremental: false
include-in-header: ../beamer_preamble.tex
bibliography: ../references.bib
fig-align: center
fig-cap-location: bottom
fig-format: pdf
links-as-notes: true
number-depth: 1
---

# Terms

target string
: The string that we want to "re-create"

input embedding
: $n_{tokens} \times n_{vocab}$ input embedding

flat input embedding
: $1 \times n_{vocab}$ input embedding

# Can we speed things up? {.allowframebreaks}

- flat input embedding of target string helps
- GPU helps a lot
- RMSProp instead of ADAM helps
  - Adjust learning rate based on input embedding and generation technique
- $n$-Beam-search is slightly slower than greedy search, but not by factor $n$ (finding better steps that take longer to compute?)

- Stop optimization based on hitting the target sequence instead? About half the iterations
- Is there a way to move to "conditional likelihoods", once we found one correct token?
- How stable is the optimization?

---

"Does it make sense to compare apples and oranges?", flat input embedding, RMSProp(lr=1e-1), T4GPU

:::{layout-ncol="2"}

![Convergence](image002.png)

![NLL](image003.png)

:::

# Generation technique?

- We run into problems if the generation technique samples: Then we cannot track the likelihood anymore (inplace gradient change)
- Likely beam search with many beams is the most balanced option?
- Contrastive search, which penalizes repeated tokens, does not work as it relies on sampling behind the scenes
- Maybe we can penalize the likelihood itself for repetitions? Or solve it through restrictions on the optimization?

# Number of parameters?

- With flat input embedding, we have $n_{vocab}$ ($768$ for GPT-2) parameters. 
- Many hyperparameters: Starting values, optimizer, generation settings, LLM
- Maybe we can reduce the number of parameters by having a smaller layer in front of the flat input-embedding. This smaller layer plus the weights connecting it to the input embedding could be less than $n_{vocab}$. (Backwards convolution?)

# Other

- Need to read up on the Google Colab resources, whether it is suitable for computing larger datasets
- GPU on Snellius (would that be feasible)?
- GitHub
- Can we compare the likelihood across different settings? (Likely not?)
- Difference to "Reverse Prompt-Engineering"