### 2024-01-18 meetings/meeting_notes.qmd

---
title: "Meeting Notes --- Summary LLM"
author: "Bas Doners, Dennis Fok, Finn-Ole Höner"
date: last-modified
date-format: long
format:
    pdf:
        documentclass: scrartcl
        toc: false
        number-sections: false
        number-depth: 2
        colorlinks: true
        fig-pos: 'H'
        fig-align: center
        geometry:
            - top = 30mm
            - bottom = 30mm
            - left = 20mm
            - right = 20mm
        papersize: dina4
aspectratio: 169
incremental: false
bibliography: ../references.bib
fig-align: center
fig-cap-location: bottom
fig-format: pdf
links-as-notes: true
number-depth: 1
---

# 18-01-2024

> First steps towards a summary LLM, basic computations seem to work.

## Can we speed things up?

## Generation technique?

## Number of parameters?

## Other

## Next steps