### 2024-01-31 meetings/meeting_notes.qmd
meeting notes

---
title: "Meeting Notes---Summary LLM"
author: "Bas Donkers, Dennis Fok, Finn-Ole Höner"
date: last-modified
date-format: long
format:
    pdf:
        documentclass: scrartcl
        toc: false
        number-sections: false
        number-depth: 2
        colorlinks: true
        fig-pos: 'H'
        fig-align: center
        geometry:
            - top = 30mm
            - bottom = 30mm
            - left = 20mm
            - right = 20mm
        papersize: dina4
bibliography: ../references.bib
fig-align: center
fig-cap-location: bottom
fig-format: pdf
links-as-notes: false
number-depth: 1
---

# 18-01-2024

- Status: First steps towards a summary LLM, basic computations seem to work.
- Admin: Setup of MS Teams group
- Technical Note: The `output` object is a tuple of length "positions", and each row of the contained matrix is one beam.

## Dutch LLMs

> I found some dutch language models on Huggingface, e.g. a [GPT-2](https://huggingface.co/GroNLP/gpt2-small-dutch) version which I use for the illustration below (also available with more parameters, see [corresponding paper](https://arxiv.org/abs/2012.05628)). There are also other “dutch” versions of other language models, see this [collection](https://huggingface.co/collections/BramVanroy/dutch-llms-64f99dddbb86c0fa80846f89). One caveat might be that all these models are in some way fine-tuned/transfer learned from the English version of the model, so not “natively” trained on only dutch corpora. However, it seems to work in our current “brute force” optimization.See attachements for the targetHeeft het zin om appels en sinaasappelen te vergelijken?? (sic)

## Speeding up computations

> Currently, we are performing too many computations. We can make use of the fact that we know the target string. Essentially, we can compute each "next word" in parallel, instead of sequentially. This also circumvents the problem that we need to think about generation strategies, as we can essentially just make a forward pass through the LLM.

## TO-DOs

- Finn speeds computations up
- Finn tries to write draft for grant proposal
- Reaching out to SKIM, slide deck, NDA

<!-- TODO: Finn speeds computations up -->
<!-- TODO: Finn tries to write draft for grant proposal -->
<!-- TODO: Reaching out to SKIM, slide deck, NDA -->