### 2024-03-14 meetings/14-03-2024_slides.qmd
typos

---
title: "Meeting --- Summary LLM"
author: "Bas Donkers, Dennis Fok, Finn-Ole Höner"
date: 2024-03-14
date-format: long
format: beamer
aspectratio: 169
incremental: false
include-in-header: ../beamer_preamble.tex
bibliography: ../references.bib
fig-align: center
fig-cap-location: bottom
fig-format: pdf
links-as-notes: true
number-depth: 1
---

# Updates

- Grant proposal handed in
- SURF Research Cloud Workshop
  - Pieter Meijndert, Yuliia Orlova
  - Small SURF Grant (quick, easy to get, uncomplicated); up to 5k GPU hours; once per calendar year
- (Start thinking about thesis comittee)
- Improving implementation of optimization still work in progress

---

{{< embed ../src/analysis/visualize_embeddings.ipynb#fig-corr1 >}}

---

{{< embed ../src/analysis/visualize_embeddings.ipynb#fig-corr2 >}}

---

{{< embed ../src/analysis/visualize_embeddings.ipynb#fig-corr3 >}}

---

{{< embed ../src/analysis/visualize_embeddings.ipynb#fig-pca >}}

---

{{< embed ../src/analysis/visualize_embeddings.ipynb#fig-tsne >}}

---

{{< embed ../src/analysis/visualize_embeddings.ipynb#fig-loo1 >}}

---

{{< embed ../src/analysis/visualize_embeddings.ipynb#fig-loo2 >}}

---

{{< embed ../src/analysis/visualize_embeddings.ipynb#fig-loo3 >}}

---

{{< embed ../src/analysis/visualize_embeddings.ipynb#fig-corr-all >}}

---

{{< embed ../src/analysis/visualize_embeddings.ipynb#fig-pca-all >}}