### 2024-04-23 meetings/12-04-2024_slides.qmd
eos works

---
title: "Meeting --- Summary LLM"
author: "Bas Donkers, Dennis Fok, Finn-Ole Höner"
date: 2024-04-12
date-format: long
format: beamer
aspectratio: 169
incremental: false
include-in-header: ../beamer_preamble.tex
bibliography: ../references.bib
fig-align: center
fig-cap-location: bottom
fig-format: pdf
links-as-notes: true
number-depth: 1
eval: false
echo: true
fig-width: 12
fig-height: 8
---

# Updates

- Improved the training, now works with tensors, GPT batching (~5-10x faster)
- Monitoring steps made is more involved than expected
- Normalization layer
- EOS token (GPT-2 has no padding token)
- "Exploration" of the encoder space
- Scales linearly with `iMaxTokens`, by the cube with `iClaims` (i.e. $O(n^3)$)

---

:::: {.columns}
::: {.column}
![Hair 10](../models/training_monitor_autoencoder.png)
:::
::: {.column}
![Hair 20](../models/training_monitor_autoencoder20.png)
:::
::::

---

# Examples of "New" Claims

Issues at $(0, 0)$?

Exact reproduction of training data, dead space, and "new" claims; i.e. EOS token and not in training data.

---

Examples:

<|endoftext|>Locks in in moisture to enhance<|endoftext|>"

<|endoftext|>Clin by enhance by enhance by saturation.<|endoftext|>

<|endoftext|>Linininineineine<|endoftext|>

<|endoftext|>Lin by enhance by enhanceistedistedistedisted by<|endoftext|>

<|endoftext|>Locks in moisture to amplify amplify l to enhance line<|endoftext|>

<|endoftext|>Elevate your confidence with confidence.<|endoftext|>

<|endoftext|>Elevate your confidence with hair that gleams in any light.<|endoftext|>

<|endoftext|>Clinically proven to enhance shine by up to%.<|endoftext|>

---

![Cleaner](./figures/cleaner_vis_norm.png)

---

![New Hair, bad grid?](./figures/output_new.png)

---

![New Hair, bad grid?](./figures/hair_new2.png)

---

![New Hair, bad grid?](./figures/hair_new3.png)

---

Inspriration from molecule generation / drug discovery [e.g @MegaSynIntegratingGenerative]

- Balancing exploration: Searching for molecules that are not only enumerations of the training data, but that are also feasible
- Primed models, focussing on specific sub-structures and saving models every $n$ epochs. This yields a spectrum of models that generate more general to more specific solutions.
- Have model that scores the quality of a claim
- Hill-climb MLE (Feedback with top X% of claims)
- Fine-tune GPT to advertising claims
- Wrap MAB around this: Play arm for new generation, evaluate with scoring model, learn from feedback (*also online on websites? This model might be rather fast? Retrain the latent space in batches based on the feedback*)
- Do we need to retrain from scratch or can we simply keep training with the existing model as a starting point?