### 2024-04-23 meetings/27-03-2024_slides.qmd

---
title: "Meeting --- Summary LLM"
author: "Bas Donkers, Dennis Fok, Finn-Ole Höner"
date: 2024-03-27
date-format: long
format: beamer
aspectratio: 169
incremental: false
include-in-header: ../beamer_preamble.tex
bibliography: ../references.bib
fig-align: center
fig-cap-location: bottom
fig-format: pdf
links-as-notes: true
number-depth: 1
eval: false
echo: true
---

# Updates

- Bias terms in encoder and decoder
- More nodes in hidden layer speeds up convergence
- Also tried out 3 nodes, results are "more noisy". However, could be interesting when claims have more "dimensions" behind them.
- Threshold: Each log-likelihood below 1 (this is softer than)

# Relevant Code

```{python}
while not bStop:
    aStore = np.zeros((1, tOHETarget.shape[0]))
    lLoss = []
    for i in range(tOHETarget.shape[0]):
        ae_outputs = ae(tOHETarget[i, :])
        outputs = lGenerate[i](ae_outputs.reshape(1, 1, -1))
        loss = llikelihood(outputs, lTargetSequence[i])

        optimizer.zero_grad()
        loss.backward()
        optimizer.step()
    bStop = np.all(aLLHistory[epoch, :] < dEps)
```

---

::: {.columns}

::: {.column}

![4:15h MacBook, Hair, 10 Claims, ADAM lr=3]("./figures/ae_training_hair_2d.png")

:::

::: {.column}
![14h MacBook, Cleaner, 10 Claims, ADAM lr=3]("./figures/training_cleaner_2d.png")

:::

:::

---

::: {.columns}

::: {.column}

![Autoencoder values of Hair 2 dimensions]("./figures/ae_2d_hair.png")

:::

::: {.column}

![Autoencoder values of Cleaner 2 dimensions]("./figures/ae_sc.png")

:::

:::

---

::: {.columns}

::: {.column}

![Embeddings AE-2D]("./figures/ae_structure_hair_2d.png")

:::

::: {.column}

![Embeddings Previously]("./figures/structure_embeddings_before.png")

:::

:::

---

Hair
```{python}
    "Experience mirror-like shine with every wash.",
    "Get dazzling, glossy hair in just one use.",
    "Transform dull strands into radiant locks.",
    "Achieve salon-quality shine at home.",
    "Enhance natural brilliance with our advanced formula.",
    "Unlock the confidence of luminous hair that turns heads.",
    "Experience the joy of hair that gleams with vitality.",
    "Embrace the radiance of hair that reflects your inner beauty.",
    "Elevate your everyday with hair that shines with sophistication.",
    "Feel the allure of irresistibly shiny hair that captivates."
```

---

::: {.columns}

::: {.column}

![Corr. Mat. 3D-AE Hair]("./figures/ae_3d_corr.png")

:::

::: {.column}

![Corr. Mat. BERT Hair]("./figures/bert_corr_hair.png")

:::

:::

---

::: {.columns}

::: {.column}

![Corr. Mat. 2D-AE Hair]("./figures/ae_corr_hair_2d.png")

:::

::: {.column}

![Corr. Mat. BERT Hair]("./figures/bert_corr_hair.png")

:::

:::

---

Surface Cleaner
```{python}
    "Our surface cleaner delivers a mirror-like shine every time.",
    "Say goodbye to streaks and smudges with our powerful formula.",
    "Achieve a noticeable sparkle on all surfaces with just one application.",
    "Experience surfaces so clean, you can see your reflection in them.",
    "Our cleaner leaves behind a gleaming, polished finish that lasts.",
    "Elevate your surroundings with a touch of brilliance.",
    "Unleash the hidden beauty of your home with our surface cleaner.",
    "Experience the joy of pristine surfaces that brighten your space.",
    "Bring a sense of clarity and freshness to your environment.",
    "Enhance the ambiance of any room with surfaces that radiate."
```

---

::: {.columns}

::: {.column}

![Corr. Mat. 2D-AE Cleaner]("./figures/ae_sc_corr.png")

:::

::: {.column}

![Corr. Mat. BERT Cleaner]("./figures/bert_sc_corr.png")

:::

:::

---

::: {.columns}

::: {.column}

![PCA 2D-AE Hair]("./figures/ae_pca_hair_2d.png")

:::

::: {.column}

![PCA BERT Hair]("./figures/bert_pca_hair.png")

:::

:::

---

::: {.columns}

::: {.column}

![PCA 2D-AE Cleaner]("./figures/ae_sc_pca.png")

:::

::: {.column}

![PCA BERT Cleaner]("./figures/bert_sc_pca.png")

:::

:::

---

::: {.columns}

::: {.column}

![LOO 2D-AE Hair]("./figures/ae_loo_hair_2d.png")

:::

::: {.column}

![LOO 2D-AE Cleaner]("./figures/ae_sc_loo.png")

:::

:::

---

![Embeddings AE-3D]("./figures/ae_structure_both.png")

---

::: {.columns}

::: {.column}

![Corr. Mat. BERT Hair]("./figures/ae_both_3d_corr.png")

:::

::: {.column}

![Corr. Mat. BERT Hair]("./figures/bert_corr_both.png")

:::

:::