### 2024-05-05 meetings/07-05-2024_slides.qmd
runs and slides

---
title: "Meeting --- Summary LLM"
author: "Bas Donkers, Dennis Fok, Finn-Ole Höner"
date: 2024-05-07
date-format: long
format: beamer
aspectratio: 169
incremental: false
include-in-header: ../beamer_preamble.tex
bibliography: ../references.bib
fig-align: center
fig-cap-location: bottom
fig-format: pdf
links-as-notes: true
number-depth: 1
eval: false
echo: true
fig-width: 12
fig-height: 8
---

# Updates

Found the source for the spikes and fixed it: `.generate()` is the wrong method, need `.forward()` $\to$ wrong conditionals! Now maximizes the likelihood and faster!

- New generated slogans seem better, space looks more interesting (have yet to do the "box" normalization)
- Maybe even further performance gains by using `past_key_values`

- Idea: LLM judge to assess grammar / quality of slogans?
- (MLFlow)

\textbf{Admin}:

- [ ] SLURM application
- [ ] Set appointment for pre-defense
- [ ] Find third assessor 
- [ ] Complete data management survey