### 2024-05-13 meetings/07-05-2024_slides.qmd
improve monitoring

---
title: "Meeting --- Summary LLM"
author: "Bas Donkers, Dennis Fok, Finn-Ole Höner"
date: 2024-05-07
date-format: long
format: beamer
aspectratio: 169
incremental: false
include-in-header: ../beamer_preamble.tex
bibliography: ../references.bib
fig-align: center
fig-cap-location: bottom
fig-format: pdf
links-as-notes: true
number-depth: 1
eval: false
echo: true
fig-width: 12
fig-height: 8
---

# Updates

Improved the optimization: `.generate()` is the wrong method, need `.forward()` $\to$ wrong conditionals! Now maximizes the likelihood and is faster!

- Space looks different now; Does maximizing the likelihood collapse the encoding space?
- Maybe even further performance gains by using `past_key_values`?

- Idea: LLM judge to assess grammar / quality of slogans?
- (MLFlow)

MPhil: Need to start working towards the paper now

---

\textbf{Admin}:

- [ ] There seem to be issues with the access to the data for me now
- [ ] SLURM application
- [ ] Set appointment for pre-defense
- [ ] Find third assessor
- [ ] Complete data management survey

---

![Maxmimum Likelihood (nLL)](./figures/07-05-24-04.png)

---

![Encoded data points](./figures/07-05-24-03.png)

---

![Exploring the encoding space with generation, x are candidates](./figures/07-05-24-01.png)

---

![Cosine similarity between the candidate points](./figures/07-05-24-02.png)