### 2024-05-13 meetings/summary.qmd

---
title: "BDS MPhil: Reverse Engineering GPT"
author: "Finn-Ole Höner; Supervisors: Bas Donkers, Dennis Fok"
date: last-modified
date-format: long
format: beamer
aspectratio: 169
incremental: false
pdf-engine: pdflatex
include-in-header: ../beamer_preamble.tex
bibliography: ../references.bib
fig-align: center
fig-cap-location: bottom
fig-format: pdf
links-as-notes: true
number-depth: 1
eval: false
echo: true
fig-width: 12
fig-height: 8
---

# Setup

**Can consumer preferences inform LLM text generation?**

- **Application:** Generating advertising claims with an LLM, based on consumer preferences.

- **Methodology:** Find input-embeddings for GPT, that generate a specific advertising claim (we call these "summary embeddings"). We model these "summary embeddings" with an Autoencoder, which gives us a "generation space" for advertising claims.

- **Toy-Example:** Tangible (e.g. "50% more visible shine after one use") and intangible (e.g. "Rediscover vibrant, joyful hair") advertising claims.

# Idea

We try to find an input-embedding that makes an LLM generate a specific output-sequence: Find the summary embedding $\boldsymbol{e^{*}}$ that maximizes the likelihood of the target-sequence $t_1, \ldots, t_L$ given the summary embedding $\boldsymbol{e}$:

$$
\boldsymbol{e^{*}} = \argmax_{\boldsymbol{e}} \sum_{i=1}^{L} \log p(t_1, \ldots, t_L | \boldsymbol{e}) \text{. }
$$

We impose restrictions on $\boldsymbol{e^{*}}$ by using an Autoencoder (AE) to model its elements. The hidden layer of this AE is our "generation space".

# Current State

- Summary embeddings generate target sequences.
- Separation of tangible and intangible claims in generation space (Figure 1).
- Grid-search exploration reveals "candidates" for new claims (red crosses) and "islands" regenerating training data claims (color-coded circles) (Figure 2).

**Next Steps:**

- Optimize optimization process.
- Enhance generation mode for clearer new claims.
- Integrate consumer preferences/ratings for generated claims.
---

::: {.columns}
::: {.column}
![Tangible and intangible claims in 2D-generation space.](./figures/07-05-24-03.png)
:::
::: {.column}
![Grid-search across 2D-generation space.](./figures/slides_pitch.png)
:::
:::