### 2024-05-15 meetings/summary.qmd
1-slide pitch

---
title: "BDS MPhil"
author: "Finn-Ole Höner; Supervisors: Bas Donkers, Dennis Fok"
date: last-modified
date-format: long
format: beamer
aspectratio: 169
incremental: false
pdf-engine: pdflatex
include-in-header: ../beamer_preamble.tex
bibliography: ../references.bib
fig-align: center
fig-cap-location: bottom
fig-format: pdf
links-as-notes: true
number-depth: 1
eval: false
echo: true
fig-width: 12
fig-height: 8
---

# Decoding Marketers' Intentions: Deterministic Generation of Ad-Claims

::: {.columns}
::: {.column}

**Setting**

- Advertising claims with meta information & ratings

**Contribution**

- Novel "summary embeddings", that make an LLM generate a specific ad-claim.

**What's done?**

- Algorithm, validation, data acquisition, Snellius application

**What's next?**

- Empirical application, first-draft

<!--

- Match and mismatch of intention and wording?

- Comparing focal brand with competitors, country, consumer ratings, brand fit, different languages, ranking of claims within set, compare branded and unbranded, sustainability messages, tangible & intangible

- Different sub-categories of hair-care e.g. intense damage, hydro, color

- mixtures of certain phrasings

Using model for prediction? Have training claims and test claims embedded in the same space. Then use the training claims to predict the test claims based on encoding space.

male and female targeting

short-term, long-term benefits

fit with brand measure

Pre-training based on GPT generated claims

Dairy: Naturality, nutrition, ingredients, ferments
-->

:::
::: {.column}
![2D-generation space.](./figures/slides_pitch.png)
:::
:::

# Setup

**Can consumer preferences inform LLM text generation?**

- **Application:** Generating advertising claims with an LLM, based on consumer preferences.

- **Methodology:** Find input-embeddings for GPT, that generate a specific advertising claim (we call these "summary embeddings"). We model these "summary embeddings" with an Autoencoder, which gives us a "generation space" for advertising claims.

- **Toy-Example:** Tangible (e.g. "50% more visible shine after one use") and intangible (e.g. "Rediscover vibrant, joyful hair") advertising claims.

# Idea

We try to find an input-embedding that makes an LLM generate a specific output-sequence: Find the summary embedding $\boldsymbol{e^{*}}$ that maximizes the likelihood of the target-sequence $t_1, \ldots, t_L$ given the summary embedding $\boldsymbol{e}$:

$$
\boldsymbol{e^{*}} = \argmax_{\boldsymbol{e}} \sum_{i=1}^{L} \log p(t_1, \ldots, t_L | \boldsymbol{e}) \text{. }
$$

We impose restrictions on $\boldsymbol{e^{*}}$ by using an Autoencoder (AE) to model its elements. The hidden layer of this AE is our "generation space".

# Current State

- Summary embeddings generate target sequences.
- Separation of tangible and intangible claims in generation space (Figure 1).
- Grid-search exploration reveals "candidates" for new claims (red crosses) and "islands" regenerating training data claims (color-coded circles) (Figure 2).

**Next Steps:**

- Optimize optimization process.
- Enhance generation mode for clearer new claims.
- Integrate consumer preferences/ratings for generated claims.

---

::: {.columns}
::: {.column}
![Tangible and intangible claims in 2D-generation space.](./figures/07-05-24-03.png)
:::
::: {.column}
![Grid-search across 2D-generation space.](./figures/slides_pitch.png)
:::
:::